---
title: "Final Project - Tools for Data Science"
author: "Sofani Mesfun"
date: "2022"
output: html_document
---

In this project you will be working with R, SQL, and Python in the same document. We will use the data sets *airlines* and *flights* from the package *nycflights13*.

1. *Knit* the empty file (to make sure everything is working)
2. Consistently *knit* the file each time you answer a question

⚠️ Students enrolled in CAP4755 should only solve: **questions 1 and 2**


**In R, install the packages ```nycflights13```, ```sqldf```, ```tydiverse``` and load all data sets and packages**.
Take your time to understand the data sets. 
```{r, warning=FALSE, message=FALSE}
#Install code here - comment the code when you finish installation
#install.packages("nycflights13")
#install.packages("sqldf")
#install.packages("tidyverse")
#install.packages("seaborn")
#install.packages("dplyr") 
# Load all packages here
library(nycflights13)
library(sqldf)
library(tidyverse)
library(ggplot2)
# Load the data here
data(nycflights13)


```


> **Question 1: List the name of airlines where the destination is ATL airport with their *average arrival delays* and sort them from the smallest to largest average arrival delays. Use data frames ```flights``` and ```airlines```.** 

We shall solve this question using R, SQL, and Python.

### R solution
You can use the R command ```inner_join()``` to join two tables.

```{r, warning=FALSE, message=FALSE}
# Code here
ListOfFlightsAirLinesATL<-data.frame(inner_join(flights, airlines, by = 'carrier')
                           %>% filter(dest=="ATL")
                           %>% group_by(name)
                           %>% drop_na()
                           %>% summarise(Ave=mean(arr_delay))
                           %>% select(name, Ave)
                           %>% arrange(desc(-Ave))
                          
              )
ListOfFlightsAirLinesATL

```


### SQL solution
Write your SQL query in the function ```sqldf()```. For exmaple,```sqldf("select * from relig_income")``` list the data frame ```relig_income```.
```{r, warning=FALSE, message=FALSE }
# Example
#sqldf("select * from relig_income limit 3")
# code here
df_flightsdata<-data.frame(flights)
df_airlinesdata <-data.frame(airlines)

sqldf("select  name, avg(arr_delay) as AVE from df_flightsdata 
       INNER JOIN df_airlinesdata
       on df_flightsdata.carrier = df_airlinesdata.carrier
       where dest = 'ATL'
       group by name
       order by AVE ASC")
```


### Python solution
Use [df.join()](https://pandas.pydata.org/docs/reference/api/pandas.DataFrame.join.html) from pandas.

```{python}
# load python libraries 
import pandas as pd
from nycflights13 import flights
from nycflights13 import airlines

# load data
df_airlines = pd.DataFrame(airlines)
df_flights = pd.DataFrame(flights)
# code here
ListOfFlightsAirLinesATLPython = df_flights.merge(df_airlines, how='inner', on='carrier')
ListOfFlightsAirLinesATLPython.query("dest == 'ATL'", inplace=True )
ListOfFlightsAirLinesATLPython.groupby(by = ['name'])['arr_delay'].mean().sort_values(ascending=True)

```



> **Question 2: Plot the boxplot of the departure delays vs the name of airlines where the destination is ATL airport. Solve this question using R and Python.**


### R solution
Use ggplot
```{r, warning=FALSE, message=FALSE}
library(dplyr, warn.conflicts = FALSE)
# Suppress summarise info
options(dplyr.summarise.inform = FALSE)
# code here
departuredelayNameofAirlineATL<-data.frame(inner_join(flights, airlines, by = 'carrier')
                           %>% filter(dest=="ATL")
                           %>% group_by(name)
                           %>% drop_na()
                           %>% summarise(dep_delay)
              )
ggplot(departuredelayNameofAirlineATL, aes(x = name, y = dep_delay,  fill = name)) + geom_boxplot() +
   stat_summary(fun = "mean", geom = "point", shape = 8,
               size = 2, color = "white")


```


### Python solution
You can use [pandas.boxplot()](https://pandas.pydata.org/docs/reference/api/pandas.DataFrame.boxplot.html) or [seaborn.boxplot()](https://seaborn.pydata.org/generated/seaborn.boxplot.html). You can either import data from the R package into Python or use the data Python package [nycflights13](https://pypi.org/project/nycflights13/) 

```{python}
# import python libraries
import matplotlib.pyplot as plt
import seaborn as sns
import pandas as pd
# data
df_airlines = pd.DataFrame(airlines)
df_flights = pd.DataFrame(flights)
departuredelayNameofAirlineATLPython= df_flights.merge(df_airlines, how='inner', on='carrier')
departuredelayNameofAirlineATLPython.query("dest == 'ATL'", inplace=True )
departuredelayNameofAirlineATLPython.groupby('name')[('dep_delay')].aggregate('min')
# Boxplot
# code here
sns.boxplot(
    data=departuredelayNameofAirlineATLPython, y="name", x="dep_delay",
     notch=True, showcaps=False,
    flierprops={"marker": "x"},
    boxprops={"facecolor": (.4, .6, .8, .5)},
    medianprops={"color": "coral"},
  )
plt.show()


```

> **Question 3: For each airlines, 1) find the month where the average departure time is the highest in the year. 2) Make a visualization to show the results. Solve this question using your preferred language R or Python.**


```{R} 
# Do not foget to specify which language are you using.

# load libraries 
library(nycflights13)
library(tidyverse)
library(dplyr, warn.conflicts = FALSE)


# Suppress summarise info
options(dplyr.summarise.inform = FALSE)
# load data
data(nycflights13)
# code here to find the months

Averagedeparturetime <- merge(airlines, flights, by ="carrier")
Averagedeparturetime %>%
  group_by(name, month) %>%
  drop_na()  %>%
  summarise(mean_dt= mean(dep_time)) %>%
  arrange(desc(mean_dt))%>%
  slice_max(mean_dt)
# code here to make the visualization

ggplot(Averagedeparturetime, aes(x = factor(month), y = dep_time,  fill = month)) + geom_boxplot() +
   stat_summary(fun = "mean", geom = "point", shape = 8,
               size = 2, color = "white")

```

